---
title: "The Informal Market as a Transfer Function"
author: "Graham Arthur"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)

library(forecast)
library(TSA)
```

# Data

```{r read-data}
e = data.table::fread("elec.csv", sep=",")
gdp = data.table::fread("gdp.csv", sep=",")
# change to ts
e = ts(e[,2], frequency = 4, start = c(1939,1))
gdp = ts(gdp[,2], frequency = 4, start = c(1947,1))


dat = ts.intersect(e, gdp)
```

# Electricity

```{r e_tsdisplay}
ggtsdisplay(dat[,1], main="Electricity Index", theme=ggplot2::theme_bw(), points=F)
```

```{r log_diff_e}
l.trend.e = lowess(log(dat[,1]), f=1/3)$y
l.diff.e = diff(log(dat[,1]), lag = 1)
ggtsdisplay(l.diff.e, main="Log Diff. Electricity", theme=ggplot2::theme_bw(), points=F)
```

# GDP

```{r gdp_tsdisplay}
ggtsdisplay(dat[,2], main="GDP", theme=ggplot2::theme_bw(), points=F)
```

```{r log_diff_gdp}
l.trend.gdp = lowess(log(dat[,2]), f=1/3)$y
l.diff.gdp = diff(log(dat[,2]), lag=1)
ggtsdisplay(l.diff.gdp, main="Log Diff. GDP", theme=ggplot2::theme_bw(), points=F)
```

# Transfer Model  {.tabset .tabset-pills}

We have to identify order processes for and GDP, and then we define the order of transfer.

## GDP Fit {.tabset}

```{r}
(model.gdp = auto.arima(l.diff.gdp, approximation=F, seasonal=T, allowdrift=F))
```

## Transfer Order

```{r ccf}
ccf(l.diff.e, l.diff.gdp, lag.max = 20, main="Elec. vs. GDP", ylab="CCF")
```

A transfer order of 2 looks good.

## Transfer Fit {.tabset}

Identify a transfer order using CCF.

```{r}
(transfer.mod = arimax(l.diff.gdp, order=arimaorder(model.gdp)[1:3],
                       seasonal=list(order=arimaorder(model.gdp)[4:6]),
                       xtransf=l.diff.e, transfer=list(c(2,0)),
                       optim.control = list(maxit = 1000), method="ML",
                       transform.pars = T))
```

### Residuals {.tabset}

```{r resid_tsdisplay}
ggtsdisplay(residuals(transfer.mod), main="Transfer Residuals", theme=ggplot2::theme_bw(), points=F)
```

#### White Noise

Are the residuals white noise?

```{r}
Box.test(residuals(transfer.mod), lag = 20, type="Ljung-Box")
```

We fail to reject that they are!

#### Gaussianity

What about gaussianity?

```{r resid_qq}
qqnorm(residuals(transfer.mod), pch=16); qqline(residuals(transfer.mod));
```

Let's do a real test.

```{r}
tseries::jarque.bera.test(residuals(transfer.mod))
```

### Fitted Values

```{r fitted_gdp}
plot_frame = data.frame(t = time(dat[,2]), 
                        gdp = dat[,2], 
                        gdp_hat =
                          dat[1,2]*c(1,cumprod(exp(fitted.values(transfer.mod))))
                        )

library(ggplot2)
ggplot(plot_frame) + 
  geom_line(aes(x=t, y=gdp, color="Actual"), size=1, alpha=0.7) +
  geom_line(aes(x=t, y=gdp_hat, color="Fitted"), size=1, alpha=0.5) +
  scale_color_manual("", values=c("black", "#419192")) +
  labs(x="Years", y="GDP (billion USD)", title="Transfer Model Fitted Values") +
  theme_bw()
```

Umm... ok cool.

# Informality {.tabset .tabset-pills}

## Absolute Informality  {.tabset}

### Raw

```{r}
informal.size = plot_frame$gdp_hat - plot_frame$gdp
plot_frame$y = informal.size
ggplot(plot_frame, aes(x = t, y = y)) +
  geom_step(size = 1.05) +
  geom_hline(yintercept = 0, linetype = 2, size = 1.1, color = "#419192") + 
  labs(x = "Years", y ="GDP, 
       title="Predicted Size of the Informal Market") +
  theme_bw()
```

### Zero-Truncated

```{r}
plot_frame$y = ifelse(informal.size < 0, 0, informal.size)
ggplot(plot_frame, aes(x = t, y = y)) +
  geom_step(size = 1.05) +
  labs(x = "Years", y ="GDP", 
       title="Predicted Size of the Informal Market") +
  theme_bw()
```

## Relative Informality {.tabset}

### Raw

```{r}
rel.informal.size = (plot_frame$gdp_hat - plot_frame$gdp)/plot_frame$gdp
plot_frame$y = rel.informal.size
ggplot(plot_frame, aes(x = t, y = y)) +
  geom_step(size = 1.05) +
  geom_hline(yintercept = 0, linetype = 2, size = 1.1, color = "#419192") + 
  labs(x = "Years", y ="Proportion of GDP", 
       title="Relative Size of the Informal Market") +
  theme_bw()
```

### Zero-Truncated

```{r}
plot_frame$y = ifelse(rel.informal.size < 0, 0, rel.informal.size)
ggplot(plot_frame, aes(x = t, y = y)) +
  geom_step(size = 1.05) +
  labs(x = "Years", y ="GD", 
       title="Relative Size of the Informal Market") +
  theme_bw()
```
