---
title: "Informality in Peru"
author: "Graham Arthur (garthur)"
date: "May 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)

library(forecast)
library(TSA)
library(dplyr)
library(ggplot2)
```

# Data

```{r read-data}
e = data.table::fread("wdi_elec.csv", sep=",", skip="Country Name", header=TRUE)
gdp = data.table::fread("wdi_gdp.csv", sep=",", skip="Country Name", header=TRUE)

# select a country of interest
target_country = "Peru"
e = e %>% filter(`Country Name` == target_country) %>% 
  select(-c(`Country Name`, `Country Code`, `Indicator Name`, `Indicator Code`)) %>%
  select_if(~sum(!is.na(.)) > 0) %>%
  tidyr::gather(key = "year", value = "kwh", convert = TRUE) 
gdp = gdp %>% filter(`Country Name` == target_country) %>% 
  select(-c(`Country Name`, `Country Code`, `Indicator Name`, `Indicator Code`)) %>%
  select_if(~sum(!is.na(.)) > 0) %>%
  tidyr::gather(key = "year", value = "gdp", convert = TRUE)


# change to ts
e = ts(e[,"kwh"], frequency = 1, start = min(e[,"year"]))
gdp = ts(gdp[,"gdp"], frequency = 1, start = min(gdp[,"year"]))

# make sure they start at the same time
dat = ts.intersect(e, gdp)
```

# Electricity

```{r e_tsdisplay}
ggtsdisplay(dat[,1], main="Electricity Index", theme=ggplot2::theme_bw(), points=F)
```

# GDP

```{r gdp_tsdisplay}
ggtsdisplay(dat[,2], main="GDP", theme=ggplot2::theme_bw(), points=F)
```

```{r log_diff_gdp}
l.gdp = log(dat[,2])
ggtsdisplay(log(dat[,2]), main="Log GDP", theme=ggplot2::theme_bw(), points=F)
```

# Transfer Model  {.tabset .tabset-pills}

We have to identify order processes for and GDP, and then we define the order of transfer.

## GDP Fit

```{r}
(model.gdp = auto.arima(dat[,2], approximation=F, seasonal=T, allowdrift=F))
```

## Electricity Fit
```{r}
(model.e = auto.arima(dat[,1], approximation=F, seasonal=T, allowdrift=F))
```

## Transfer Order

```{r ccf}
ccf(residuals(model.e), residuals(model.gdp), lag.max = 20, main="Elec. vs. GDP", ylab="CCF")
```

A transfer order of at most 1 looks okay.

## Transfer Fit {.tabset}

Identify a transfer order using CCF.

```{r}
(transfer.mod = arimax(dat[,2], order=arimaorder(model.gdp),
                       xtransf=dat[,1], transfer=list(c(1,0)),
                       optim.control = list(maxit = 1000), method="ML",
                       transform.pars=TRUE))
```

### Residuals {.tabset}

```{r resid_tsdisplay}
ggtsdisplay(residuals(transfer.mod), main="Transfer Residuals", theme=ggplot2::theme_bw(), points=F)
```

#### White Noise

```{r}
Box.test(residuals(transfer.mod), lag = 20, type="Ljung-Box")
```

#### Gaussianity

```{r resid_qq}
qqnorm(residuals(transfer.mod), pch=16); qqline(residuals(transfer.mod));
```

Let's do a real test.

```{r}
tseries::jarque.bera.test(residuals(transfer.mod))
```

### Fitted Values

```{r fitted_gdp}
plot_frame = data.frame(t = time(dat[,2]), 
                        gdp = dat[,2], 
                        gdp_hat = fitted.values(transfer.mod)
)

library(ggplot2)
ggplot(plot_frame) + 
  geom_line(aes(x=t, y=gdp, color="Actual"), size=1, alpha=0.7) +
  geom_line(aes(x=t, y=gdp_hat, color="Fitted"), size=1, alpha=0.5) +
  scale_color_manual("", values=c("black", "#419192")) +
  labs(x="Years", y="GDP per Capita", title="Transfer Model Fitted Values") +
  theme_bw()
```

# Informality {.tabset .tabset-pills}

## Absolute Informality  {.tabset}

### Raw

```{r}
informal.size = plot_frame$gdp_hat - plot_frame$gdp
plot_frame$y = informal.size
ggplot(plot_frame, aes(x = t, y = y)) +
  geom_step(size = 1.05) +
  geom_hline(yintercept = 0, linetype = 2, size = 1.1, color = "#419192") + 
  labs(x = "Years", y ="GDP Per Capita", 
       title="Predicted Size of the Informal Market") +
  theme_bw()
```

### Zero-Truncated

```{r}
plot_frame$y = ifelse(informal.size < 0, 0, informal.size)
ggplot(plot_frame, aes(x = t, y = y)) +
  geom_step(size = 1.05) +
  labs(x = "Years", y ="GDP Per Capita", 
       title="Predicted Size of the Informal Market") +
  theme_bw()
```

## Relative Informality {.tabset}

### Raw

```{r}
rel.informal.size = (plot_frame$gdp_hat - plot_frame$gdp)/plot_frame$gdp
plot_frame$y = rel.informal.size
ggplot(plot_frame, aes(x = t, y = y)) +
  geom_step(size = 1.05) +
  geom_hline(yintercept = 0, linetype = 2, size = 1.1, color = "#419192") + 
  labs(x = "Years", y ="Proportion of GDP per Capita", 
       title="Relative Size of the Informal Market") +
  theme_bw()
```

### Zero-Truncated

```{r}
plot_frame$y = ifelse(rel.informal.size < 0, 0, rel.informal.size)
ggplot(plot_frame, aes(x = t, y = y)) +
  geom_step(size = 1.05) +
  labs(x = "Years", y ="GDP Per Capita", 
       title="Relative Size of the Informal Market") +
  theme_bw()
```



